---
title: "Skeena and Nass Chinook Salmon Spawner-recruit Analysis"
date: '`r format(Sys.time(), "%d %B, %Y")`'
author: ""
output:
  html_document: 
    fig_caption: true
    toc: true
    toc_depth: 2
    toc_float: true
---

<style type="text/css">
.caption {
    font-size: large;
}
</style>


```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(gsl)
library(ggcorrplot)
library(ggpubr)
library(viridis)
knitr::opts_chunk$set(echo = FALSE,warning = FALSE, message = FALSE, 
                      fig.width = 6,fig.height = 4, dpi = 300)
```


```{r read and wrangle data}
# read data
source(here("R/SR_posterior_processing.R")) #can take some time
skeena_order <- c("Lower Skeena", "Kitsumkalum", "Zymoetz", "Middle Skeena", "Large Lakes", 
                  "Upper Skeena", "Skeena Aggregate")
nass_order <- c("Lower Nass", "Middle Nass", "Upper Nass", "Nass Aggregate")
```

State-space spawner-recruit models were fit [here](https://github.com/Pacific-salmon-assess/skeena-nass-chinook-goals/blob/main/R/SRR_fit.R), and model validation can be found [here](https://github.com/Pacific-salmon-assess/skeena-nass-chinook-goals/blob/main/R/SR_diagnostics.Rmd). Below we explore the output from these models. 

# Skeena  

## Data   
```{r sp_har}
sp_har_fig <- sp_har |>
  group_by(CU, year) |>
  #calculate random observation error, 4000 = iter
  mutate(spwn_LCI = quantile(rnorm(4000, spwn, spwn_cv*spwn), 0.25),  
         spwn_UCI = quantile(rnorm(4000, spwn, spwn_cv*spwn), 0.75), 
         harv_LCI = quantile(rnorm(4000, harv, harv_cv*harv), 0.25),  
         harv_UCI = quantile(rnorm(4000, harv, harv_cv*harv), 0.75))

ggplot(filter(sp_har_fig, SMU == "Skeena"), aes(year, spwn/1000)) +
  geom_ribbon(aes(xintercept = year, ymin = spwn_LCI/1000, ymax = spwn_UCI/1000), 
                width = 0, fill = "grey") +
  geom_line() +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free_y") +
  theme_sleek() +
  labs(x = "Return year", y = "Spawners (000s) and 50th percentile")

ggplot(filter(sp_har_fig, SMU == "Skeena"), aes(year, harv/1000)) +
  geom_ribbon(aes(xintercept = year, ymin = harv_LCI/1000, ymax = harv_UCI/1000), 
                width = 0, fill = "grey") +
  geom_line() +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free_y") +
  theme_sleek() +
  labs(x = "Return year", y = "Harvest (000s) and 50th percentile")
```

```{r age props}
prop_age |>
  pivot_longer(4:6, names_to = "age") |>
  filter(SMU == "Skeena") |>
  mutate(age = gsub("a", "", age)) |>
ggplot(aes(year, value, fill = age)) +
  geom_col(position = position_stack(reverse = TRUE), width = 1) +
  scale_fill_viridis_d() +
  facet_wrap(~factor(CU, levels = skeena_order)) +
  theme_sleek() +
  labs(x = "Return year", y = "Proportion-at-age", fill = "Age")
```

## Priors  

Here, we look specifically at the $S_{Max}$ prior, as some may consider it "informative".
This prior is a normal distribution with a mean of half the maximum observed spawners, and a standard deviation of 1.5 times the maximum observed spawners.  

```{r Smax_prior}
S_max_plot_data <- sp_har |>
  filter(SMU == "Skeena") |>
  group_by(CU) |>
  summarise(rnorm(nrow(filter(par.posts, SMU == "Skeena"))/length(unique(filter(sp_har, SMU == "Skeena")$CU)), 
                  max(spwn)*.5, max(spwn)*1.5)) |> #needs to match prior in stan model
  cbind(filter(par.posts, SMU == "Skeena")$S_max) 
  
colnames(S_max_plot_data) <- c("CU", "prior", "posterior")

S_max_plot_data <- S_max_plot_data |>
  pivot_longer(-CU, names_to = "par")|>
  filter(value>0) |>
  group_by(CU) |>
  filter(value < quantile(value, 0.99))

ggplot(S_max_plot_data, aes(value/1000)) +
  geom_density(aes(fill = par, color = par), alpha = 0.5) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free") +
  labs(x = "S_Max (000s)", fill = "", color = "") +
  theme_sleek()

```

## Spawner-recruit fits 

```{r static fit}
# SR fits ----
ggplot() +
  geom_abline(intercept = 0, slope = 1,col="dark grey") +
  geom_ribbon(data = filter(SR.preds, SMU == "Skeena"), 
              aes(x = Spawn/1000, ymin = Rec_lwr/1000, ymax = Rec_upr/1000),
              fill = "grey80", alpha=0.5, linetype=2, colour="gray46") +
  geom_errorbar(data = filter(brood.all, SMU == "Skeena"), 
                aes(x= S_med/1000, y = R_med/1000, ymin = R_lwr/1000, ymax = R_upr/1000),
                colour="grey", width=0, linewidth=0.3) +
  geom_errorbar(data = filter(brood.all, SMU == "Skeena"), 
                 aes(y = R_med/1000, xmin = S_lwr/1000, xmax = S_upr/1000),
                 height=0, colour = "grey", linewidth = 0.3) +
  geom_point(data = filter(brood.all, SMU == "Skeena"),
             aes(x = S_med/1000, y = R_med/1000, color=BroodYear), size = 1.5) +
  geom_line(data = filter(SR.preds, SMU == "Skeena"), 
            aes(x = Spawn/1000, y = Rec_med/1000)) +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free") +
  scale_colour_viridis_c(name = "Brood Year")+
  labs(x = "Spawners (000s)", y = "Recruits (000s)") +
  theme_sleek()
```

Latent (i.e., hidden) states of spawners and recruits fir in the spawner-recruit model  
```{r latent states}
ggplot(filter(AR1.spwn, SMU == "Skeena"), aes(year, S.50/1000)) +
  geom_ribbon(aes(ymin = S.25/1000, ymax = S.75/1000), 
              fill = "grey") +
  geom_line() +
  geom_point(data = filter(sp_har, SMU == "Skeena"), aes(year, spwn/1000)) +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free_y") +
  theme_sleek() +
  labs(x = "Return year", y = "Spawners (000s) and 50th percentile")

ggplot(filter(AR1.harv, SMU == "Skeena"), aes(year, H.50/1000)) +
  geom_ribbon(aes(ymin = H.25/1000, ymax = H.75/1000), 
              fill = "grey") +
  geom_line() +
  geom_point(data = filter(sp_har, SMU == "Skeena"), aes(year, harv/1000)) +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free_y") +
  theme_sleek() +
  labs(x = "Return year", y = "Harvest (000s) and 50th percentile")
```


And a fit with time-varying productivity (Ricker $\alpha$):

```{r tv alpha fit}
a.yrs.all |>
  filter(SMU == "Skeena", 
         brood_year < max(a.yrs.all$brood_year)-a_max) |> # index fully observed rec
  ggplot(aes(color=CU)) +
  geom_line(aes(x = brood_year , y = mid), lwd = 1.5) +
  scale_color_viridis_d() +
  geom_hline(yintercept = 1, lty=2, col = "grey") +
  labs(y ="Productivity (\U03B1)", x = "Brood year")+
  guides(color=guide_legend(title="Stock")) +
  scale_y_continuous(limits = c(0,8)) +
  annotate("text", x = 1990, y = 1.5, label = "replacement", color = "grey") +
  theme_sleek() +
  theme(legend.position = "bottom") 
  
a.yrs.all |>
  filter(SMU == "Skeena", 
         brood_year < max(a.yrs.all$brood_year)-a_max) |> # index fully observed rec
  ggplot() +
  geom_ribbon(aes(x = brood_year, ymin = lwr, ymax = upr), fill = "grey") +
  geom_line(aes(x = brood_year , y = mid)) +
  geom_hline(yintercept = 1, lty=2, col = "grey") +
  labs(y ="Productivity (\U03B1) and 80th percentile", x = "Brood year")+
  facet_wrap(~factor(CU, levels = skeena_order)) +
  scale_y_continuous(limits = c(0,13)) +
  theme_sleek() +
  theme(legend.position = "bottom") 

ggplot() +
  geom_point(data = filter(brood.all, SMU == "Skeena"),
             aes(x = S_med/1000,
                 y = R_med/1000),
             size = 1.5) +
  geom_line(data = filter(TV.SR.preds, SMU == "Skeena"), aes(x = spw/1000, y = pred.R/1000, color = year, group = year)) +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free") +
  scale_colour_viridis_c(name = "Brood Year")+
  geom_abline(intercept = 0, slope = 1,col="dark grey") +
  labs(x = "Spawners (000s)",
       y = "Recruits (000s)",
       title = "Time varying productivity spawner-recruit fits") +
  theme_sleek()
```

We also check residuals from both models:

```{r residuals}
# AR1 resids ----
ggplot(filter(AR1.resids, SMU == "Skeena"), aes(x=year, y = mid)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr),  fill = "darkgrey", alpha = 0.5) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupr),  fill = "black", alpha=0.2) +
  geom_line(lwd = 1.1) +
  coord_cartesian(ylim=c(-2,2)) +
  labs(x = "Return year",
       y = "Recruitment residuals",
       title = "AR1 recruitment residuals") +
  facet_wrap(~factor(CU, levels = skeena_order)) +
  theme(legend.position = "none",
        panel.grid = element_blank()) +
  theme_sleek() +
  geom_abline(intercept = 0, slope = 0, col = "dark grey", lty = 2)


# TV resids ----
ggplot(filter(TV.resids, SMU == "Skeena"), aes(x=year, y = mid)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr),  fill = "darkgrey", alpha = 0.5) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupr),  fill = "black", alpha=0.2) +
  geom_line(lwd = 1.1) +
  labs(x = "Return year",
       y = "Recruitment residuals",
       title = "Time-varying recruitment residuals") +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free_y") +
  theme(legend.position = "none",
        panel.grid = element_blank()) +
  theme_sleek() +
  geom_abline(intercept = 0, slope = 0, col = "dark grey", lty = 2)
```

```{r benchmarks}
bench.posts |>
  filter(SMU == "Skeena") |>
  select(CU, Smsy, Umsy) |>
  cbind(ln_a = filter(par.posts, SMU == "Skeena")$ln_a) |>
  pivot_longer(-1, names_to = "par") |>
  group_by(CU) |>
  filter(value < quantile(value, 0.90)) |>
  ggplot(aes(value, fill = CU, color = CU)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~par, nrow = 3, scales = "free", strip.position = "bottom") +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  theme_sleek() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), 
        strip.background = element_blank(),
        strip.placement = "outside") +
  labs(x = "", y = "")
```

```{r spawner status}
bench_plot <- filter(bench.par.table, SMU == "Skeena", 
                     par %in% c("Sgen", "Smsy.80")) |>
  select(CU, par, median)

ggplot(filter(AR1.spwn, SMU == "Skeena"), aes(year, S.50/1000)) +
  geom_line() +
  geom_ribbon(aes(ymin = S.25/1000, ymax = S.75/1000), fill = "darkgrey", alpha = 0.5) +
  geom_hline(data = bench_plot, 
             aes(yintercept = median/1000, col = par), lty = "dashed") +
    scale_colour_manual(values=c(Smsy.80 = "forestgreen", 
                                 Sgen = "darkred"),
                      labels=c(Smsy.80 = expression(paste("80%", S[MSY])),
                               Sgen = expression(S[Gen]))) +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free_y") +
  theme_sleek() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(x = "Retun Year", y = "Spawners (000s)", col = "")
```

```{r bench-posts}
bench.long <- select(bench.posts, CU, SMU, Sgen, Smsy.80, S.recent) |>
  pivot_longer(cols = c(Sgen, Smsy.80, S.recent),
                           names_to = "par") |>  
  arrange(CU, par, value) |>
  group_by(CU, par) |>
  filter(between(value, quantile(value, 0.05), quantile(value, 0.95)), 
         SMU == "Skeena") 

ggplot(bench.long, aes(value/1000, fill = par, color = par)) +
  geom_density(alpha = 0.3) +
  geom_vline(xintercept = 1.5, lty=2, col="grey") +
  facet_wrap(~factor(CU, levels = skeena_order), scales = "free") +
  scale_fill_manual(breaks = c("S.recent", "Sgen", "Smsy.80"),
                    values = c("black", "darkred", "forestgreen"),
                    aesthetics = c("fill", "color"),
                    labels = c(expression(italic(S[recent])), 
                               expression(italic(S[Gen])),
                               expression(italic(paste("80% ",S)[MSY])))) +
  labs(x = "Spawners (000s)", y = "Posterior density", fill = "", color = "") +
  theme_sleek()
```

```{r bench compare, width=700*2, height=900*2, dpi=240}
bench_compare <- bench.par.table |>
  filter(par %in% c("Sgen", "Smsr.20", "Smsy.80", "Smsr.40", "Seq", "Smsr"), 
         SMU == "Skeena", 
         CU != "Skeena Aggregate")

a <- ggplot(filter(bench_compare, par %in% c("Sgen", "Smsr.20")), 
            aes(x = CU, y = median/1000, fill = par)) +
  geom_bar(position="dodge", stat = "identity") +
  geom_errorbar(aes(ymin = lwr/1000, ymax = upr/1000 ,col = par), 
                width = 0,position=position_dodge(0.9)) +
  theme_sleek() +
  scale_fill_manual(values = scales::hue_pal()(2),
                    labels=c(expression("S"["Gen"]), expression("20%S"["MSR"])),
                    aesthetics=c("fill", "colour")) +
  theme(legend.position = c(0.8,0.7),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.title=element_blank()) +
  labs(x = "Conservation Unit", y = "")

b <- ggplot(filter(bench_compare, par %in% c("Smsy.80", "Smsr.40")), 
            aes(x = CU, y = median/1000, fill = par)) +
  geom_bar(position="dodge", stat = "identity") +
  geom_errorbar(aes(ymin = lwr/1000, ymax = upr/1000, col = par), 
                width = 0, position = position_dodge(0.9)) +
  theme_sleek()  +
  scale_fill_manual(values = scales::hue_pal()(2),
                    labels=c(expression("80%S"["MSY"]), expression("40%S"["MSR"])),
                    aesthetics=c("fill", "colour")) +
  theme(legend.position = c(0.8,0.7),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.title=element_blank()) +
  labs(x = "Conservation Unit", y = "")

c <- ggplot(filter(bench_compare, par %in% c("Seq", "Smsr")),
            aes(x = CU, y = median/1000, fill = par)) +
  geom_bar(position="dodge", stat = "identity") +
  geom_errorbar(aes(ymin = lwr/1000, ymax = upr/1000, col = par), 
                width = 0, position = position_dodge(0.9)) +
  scale_fill_manual(values = scales::hue_pal()(2),
                    labels=c(expression("S"["eq"]), expression("S"["MSR"])),
                    aesthetics=c("fill", "colour")) +
  theme_sleek() +
  theme(legend.position = c(0.8,0.7),
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        legend.title=element_blank()) +
  labs(x = "Conservation Unit", y = "")

cowplot::plot_grid(a, b, c, labels="auto", ncol=1,rel_heights=c(0.6,0.6,1))


```

## Demographics 

```{r demography}
#this may belong in another file somewhere 


```

# Nass  